---
title: "`r params$opt$project`" 
date: "`r paste('Date:', format(Sys.time(), '%d %B, %Y'))`"
vignette: >
  %\VignetteIndexEntry{CellBridge report}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
output:
  html_document:
    fig_caption: true
    df_print: kable
    toc: true
    toc_float: true
params:
    obj_ls: NULL
    scrub_ls: NULL
    meta_data: NULL
    pre.qc.summ.rna: NULL
    opt: NULL
    ui: NULL
---

<style> 
  #header { 
    background: -moz-linear-gradient(left, rgba(111,19,221,1) 0%, rgba(125,185,232,0) 75%); /* FF3.6-15 */
    ackground: -webkit-linear-gradient(left, rgba(111,19,221,1) 0%,rgba(125,185,232,0) 75%); /* Chrome10-25,Safari5.1-6 */
    background: linear-gradient(to right, rgba(111,19,221,1) 0%,rgba(125,185,232,0) 75%); /* W3C, IE10+, FF16+, Chrome26+, Opera12+, Safari7+ */
    filter: progid:DXImageTransform.Microsoft.gradient( startColorstr='#6f13dd', endColorstr='#007db9e8',GradientType=1 ); /* IE6-9 */
    color: white;
    height: 110px;
  }
</style>


```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri("cellbridge_logo.png"), 
               alt = 'logo', 
               style = 'position:absolute; top:22px; right:40px; width: 2.0in'
               )
```


```{r variable, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
    obj_ls <- params$obj_ls
    scrub_ls <- params$scrub_ls
    meta_data <- params$meta_data
    pre.qc.summ.rna <- params$pre.qc.summ.rna
    opt <- params$opt
    ui <- params$ui
    n.sample <- length(obj_ls)
```

```{r, include=FALSE}
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```

- - -

# Parameters used in the QC overview 
Note: no filter is applied in this summary.

```{r table-1, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
kp.mtdata <- c("species", "tissue", "metadata", "min_umi_per_cell", "max_umi_per_cell",
               "min_genes_per_cell", "max_genes_per_cell", "max_mt_percent", "min_cell", "scr_th")
data.frame(opt) %>% 
  dplyr::select(names(opt)[names(opt) %in% kp.mtdata]) %>%
  dplyr::select(all_of(kp.mtdata[kp.mtdata %in% names(.)])) %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "110px")
```

- - -

# Quality control

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
umi_per_cell <- data.frame()
gene_per_cell <- data.frame()
mt_per_cell <- data.frame()
for (i in 1:length(obj_ls)) {
  exprMat <- obj_ls[[i]]
  smpl <-  names(obj_ls)[i]
  # UMI per cell
  countByCell <- colSums(exprMat)
  summ1 <- c(min = min(countByCell), 
             quantile(countByCell, c(0.0, 0.25, 0.5, 0.75, 1)),
             max = max(countByCell))
  umi_per_cell <- umi_per_cell %>%
    dplyr::bind_rows(scales::comma(summ1, accuracy = 1))
  # Gene per cell
  countByCell <- colSums(exprMat > 0)
  summ2 <- c(min = min(countByCell), 
             quantile(countByCell, c(0.0, 0.25, 0.5, 0.75, 1)),
             max = max(countByCell))
  gene_per_cell <- gene_per_cell %>%
    dplyr::bind_rows(scales::comma(summ2, accuracy = 1))
  # Mitochondrial (%) per cell
  # creat seurat object
  sobj <- CreateSeuratObject(counts=obj_ls[[i]], project=smpl, 
                             min.cells=0, min.features=0)
  # mitochondrial genes
  if (opt$species == "hs") {
    sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = "^MT-")  
    }
  if (opt$species == "mm") {
    sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = "^mt-")  
  }
  if (opt$species == "mf") {
    mito_genes <- c('ND5','COX3','ATP8','COX1','ND6','ND3','ND4L','COX2','ND1','CYTB','ATP6','ND4','ND2')
    sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = mito_genes)  
  }
  pctmt <- sobj$percent_mt
  summ3 <- c(min = min(pctmt), 
             quantile(pctmt, c(0.0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
             max = max(pctmt))
  mt_per_cell <- mt_per_cell %>%
    dplyr::bind_rows(round(summ3, 2))
}
rownames(umi_per_cell) <- names(obj_ls)
rownames(gene_per_cell) <- names(obj_ls)
rownames(mt_per_cell) <- names(obj_ls)
```

### Each sample characteristics, demographics, and QC metrics {.tabset}

#### Metadata

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
meta_data %>%
  dplyr::mutate(sample_id = factor(sample_id, levels = sample_id[mixedorder(sample_id)], ordered = TRUE)) %>%
  dplyr::arrange(sample_id) %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "200px")
```

#### Cells & Genes

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
pre.qc.summ.rna %>%
  dplyr::rename(pre_qc_cell = cell,
                pre_qc_gene = gene) %>%
  dplyr::mutate(pre_qc_cell = scales::comma(pre_qc_cell, accuracy = 1),
         pre_qc_gene = scales::comma(pre_qc_gene, accuracy = 1)) %>%
  dplyr::select(-type) %>%
  dplyr::mutate(sample_id = factor(sample_id, levels = sample_id[mixedorder(sample_id)], ordered = TRUE)) %>%
  dplyr::arrange(sample_id) %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "200px")
```

#### UMI per cell

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
umi_per_cell %>%
  tibble::rownames_to_column(var="sample_id") %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "200px")
```

#### Gene per cell

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
gene_per_cell %>%
  tibble::rownames_to_column(var="sample_id") %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "200px")
```

#### Mitochondrial (%) per cell

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.align = 'center'}
mt_per_cell %>%
  tibble::rownames_to_column(var="sample_id") %>%
  kable() %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, fixed_thead = TRUE) %>%
  scroll_box(width = "100%", height = "200px")
```

### {-}

- - -

```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.width=17.5, fig.height=4.5, fig.align = 'center'}
  # ===================================
  set.seed(123)  
  colrs <- color_helper(set.names=sort(unique(meta_data$sample_id)), panel="Set1") 
  colrs <- setNames(as.vector(colrs)[sample(1:length(colrs), size = length(colrs), replace = FALSE,)], names(colrs)) 
  # ===================================
  db_plot <- data.frame()
  for (i in 1:length(obj_ls)) {
    obj <- obj_ls[[i]]
    smpl <-  names(obj_ls)[i]
    # creat seurat object
    sobj <- CreateSeuratObject(counts=obj, project=smpl, 
                               min.cells=0, min.features=0)
    # mitochondrial genes
    if (opt$species == "hs") {
      sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = "^MT-")  
      }
    if (opt$species == "mm") {
      sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = "^mt-")  
    }
    if (opt$species == "mf") {
      mito_genes <- c('ND5','COX3','ATP8','COX1','ND6','ND3','ND4L','COX2','ND1','CYTB','ATP6','ND4','ND2')
      sobj[["percent_mt"]] <- PercentageFeatureSet(sobj, pattern = mito_genes)  
      }
    # head(sobj@meta.data)
    db_plot <- db_plot %>%
      dplyr::bind_rows(sobj@meta.data %>%
                         dplyr::mutate(barcode = rownames(.)) %>%
                         dplyr::select(barcode, nCount_RNA, nFeature_RNA, percent_mt) %>%
                         tidyr::gather(., key = type, value = count, -barcode)%>%
                         dplyr::mutate(sample_id = smpl))
  }
  # ===================================
filters <- data.frame(threshold = c(opt$min_umi_per_cell, 
                                    ifelse(is.null(opt$max_umi_per_cell), 0, opt$max_umi_per_cell), 
                                    opt$min_genes_per_cell, 
                                    ifelse(is.null(opt$max_genes_per_cell), 0, opt$max_genes_per_cell), 
                                    opt$max_mt_percent),
                      type = c("nCount_RNA", "nCount_RNA", 
                               "nFeature_RNA", "nFeature_RNA", 
                               "percent_mt"))
  # ===================================
  Q <- list()
  for (tp in unique(db_plot$type)) {
    p <- ggplot() +
      getBaseTheme() +
      theme(strip.text = element_text(size = 20, face = "bold"),
            axis.title = element_text(size = 14, face = "bold"),
            axis.text = element_text(size = 12, face = "bold"),
            legend.title = element_text(size=13),
            legend.text = element_text(size=11)) +
      scale_x_continuous(trans = scales::log10_trans(),
                         breaks = scales::trans_breaks("log10", function(x) 10^x),
                         labels = scales::trans_format("log10", scales::math_format(10^.x))) +
      xlab("10^x") + ylab("Density") +
      geom_density(data = filter(db_plot, type == tp),
                   aes(x = feature, color = sample_id), 
                   show.legend = FALSE, na.rm = TRUE) +
      geom_vline(data = filter(filters, type == tp),
                 aes(xintercept = threshold),
                 linetype = 2, color = "firebrick", size = 0.45, na.rm = TRUE) +
      scale_color_manual(values = colrs) +
      guides(fill = "none", color = "none")
    Q[[length(Q) + 1]] <- p  
  }
  # ===================================
```

### Distribution Plots for Quality Control Metrics {.tabset}

Dashed lines indicate the thresholds provided to the pipeline.

#### UMI per cell

<center>
```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(5.5, 4.5), fig.align = 'center'}
ggplotly(Q[[1]], tooltip = c("sample_id"))
```
<center>

### Gene per cell

<center>
```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(5.5, 4.5), fig.align = 'center'}
ggplotly(Q[[2]], tooltip = c("sample_id"))
```
<center>

### Mitochondrial (%) per cell

<center>
```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(5.5, 4.5), fig.align = 'center'}
ggplotly(Q[[3]], tooltip = c("sample_id"))
```
<center>

### {-}

- - -

```{r, eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
 # Check overlap of cell barcodes
samples <- unique(meta_data$sample_id)

bc.qc.raw <- matrix(nrow = n.sample, ncol = n.sample)
colnames(bc.qc.raw) <- as.vector(samples)
rownames(bc.qc.raw) <- as.vector(samples)

bc.qc.ji <- matrix(nrow = n.sample, ncol = length(samples))
colnames(bc.qc.ji) <- as.vector(samples)
rownames(bc.qc.ji) <- as.vector(samples)

for(i in samples){
  for(j in samples){
    qi <- obj_ls[[i]]
    qj <- obj_ls[[j]]
    qc <- length(intersect(colnames(qi), colnames(qj)))
    bc.qc.raw[i,j] <- qc
    bc.qc.ji[i,j] <- qc/mean(c(dim(qi)[2],dim(qj)[2]))
  }
}

bc.qc.raw <- bc.qc.raw[!apply(bc.qc.raw, 1, function(x) sum(x == 0) == (n.sample-1)), 
                       !apply(bc.qc.raw, 2, function(x) sum(x == 0) == (n.sample-1))]

bc.qc.ji <- round(bc.qc.ji*100, 2)
bc.qc.ji <- bc.qc.ji[!apply(bc.qc.ji, 1, function(x) sum(x == 0) == (n.sample-1)), 
                     !apply(bc.qc.ji, 2, function(x) sum(x == 0) == (n.sample-1))]
```

### Barcodes contamination {.tabset}

#### Raw count

Number of barcodes shared between pairs of samples.

```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(5.5, 5), fig.align = 'center'}
if (length(bc.qc.raw) != 0) {
  bc.qc.raw %>%
    kable() %>%
    kable_classic(full_width = F, html_font = "Cambria") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                  full_width = F, fixed_thead = TRUE) %>%
    scroll_box(width = "100%", height = "200px")  
} else {
  paste("No overlaps among samples' barcodes.")
}
```

#### Jaccard Index

Fraction (%) of barcodes shared between pairs of samples.

```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(5.5, 5), fig.align = 'center'}
if (length(bc.qc.ji) != 0) {
  bc.qc.ji %>%
    kable() %>%
    kable_classic(full_width = F, html_font = "Cambria") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                  full_width = F, fixed_thead = TRUE) %>%
    scroll_box(width = "100%", height = "200px")  
} else {
  paste("No overlaps among samples' barcodes.")
}
```

### {-}

- - -

`r keyval <- opt$scr_th != 0`

`r if (keyval) {"### Doublet scores distribution {.tabset}"}`

```{r, eval=keyval, warning=FALSE, message=FALSE, echo=FALSE}
# plot scrublet scores
Q <- list()

dp <- lapply(scrub_ls, function(x){
  data.frame(x["doublet_scores_obs"])
  })
dp <- map_df(dp, ~as.data.frame(.x), .id="id") %>%
  dplyr::rename(sample_id = id,
                score = doublet_scores_obs) %>%
  dplyr::mutate(type = "Observed")
p <- ggplot() +
  getBaseTheme() +
  theme(strip.text = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        legend.title = element_text(size=9),
        legend.text = element_text(size=8)) +
  xlab("Doublet score") + ylab("Density") +
  geom_density(data = dp,
               aes(x = score, group = sample_id, color = sample_id), show.legend = FALSE) +
  scale_color_manual(name = "Sample",
                     values = colrs) +
  guides(color = guide_legend(nrow = ifelse(n.sample > 4, 2, 1))) + 
  geom_vline(xintercept=opt$scr_th, linetype=2, color="firebrick", size=0.45)
Q[[length(Q) + 1]] <- p
  
dp <- lapply(scrub_ls, function(x){
  data.frame(x["doublet_scores_sim"])
  })
dp <- map_df(dp, ~as.data.frame(.x), .id="id") %>%
  dplyr::rename(sample_id = id,
                score = doublet_scores_sim) %>%
  dplyr::mutate(type = "Simulated")
p <- ggplot() +
  getBaseTheme() +
  theme(strip.text = element_text(size = 14, face = "bold"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12, face = "bold"),
        legend.title = element_text(size=9),
        legend.text = element_text(size=8)) +
  xlab("Doublet score") + ylab("Density") +
  geom_density(data = dp,
               aes(x = score, group = sample_id, color = sample_id), show.legend = FALSE) +
  scale_color_manual(name = "Sample",
                     values = colrs) + 
  geom_vline(xintercept=opt$scr_th, linetype=2, color="firebrick", size=0.45)
Q[[length(Q) + 1]] <- p
```

`r if (keyval) {"#### Observed scores"}`

```{r eval=keyval, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(4.0, 3.0), fig.align = 'center'}
plot(Q[[1]])
```

`r if (keyval) {"#### Simulated scores"}`

```{r eval=keyval, warning=FALSE, message=FALSE, echo=FALSE, fig.dim=c(4.0, 3.0), fig.align = 'center'}
plot(Q[[2]])
```

`r if (keyval) {"### {-}"}`

`r if (keyval) {"- - -"}`

- - -

# Session information

This is the output of `sessionInfo()` on the system on which this document was compiled

```{r eval=TRUE, warning=FALSE, message=FALSE, echo=FALSE, max.height='300px'}
sessionInfo()
```
